---
title: "Getting Started with PurpleAirAPI"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with PurpleAirAPI}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup}
library(PurpleAirAPI)
```

## Introduction

The PurpleAirAPI package provides tools for accessing air quality data from PurpleAir sensors. This vignette demonstrates how to:

1. Get sensor metadata
2. Download historical data

## Getting Sensor Information

First, let's get information about available sensors:

```{r, eval=FALSE}
# Replace with your API key
api_key <- "YOUR_API_KEY"
```

```{r, eval=FALSE}
# Get sensor locations
sensors <- getPurpleairSensors(
  apiReadKey = api_key,
  fields = c("latitude", "longitude", "date_created", "last_seen", "location_type")
)
```

```{r, include=FALSE}
# Example output structure (not real data):
sensors <- data.frame(
  sensor_index = c(1234, 5678),
  latitude = c(37.7749, 34.0522),
  longitude = c(-122.4194, -118.2437),
  date_created = as.Date(c("2020-01-01", "2020-02-01")),
  last_seen = as.Date(c("2024-01-01", "2024-01-01")),
  location_type = c(0, 1),
  stringsAsFactors = FALSE
)
```

```{r, eval=TRUE}
print(sensors)
```

## Downloading Historical Data

Once you have identified sensors of interest, you can download their historical data:

```{r, eval=FALSE}
# Download one week of hourly data
history <- getSensorHistory(
  sensorIndex = c(1234, 5678),  # Example sensor IDs
  apiReadKey = api_key,
  startDate = "2024-01-01",
  endDate = "2024-01-07",
  average = 60,  # hourly averages
  fields = c("pm2.5_alt", "temperature", "humidity")
)
```

```{r, include=FALSE}
# Example output structure (not real data):
history <- data.frame(
  time_stamp = c("2024-01-01 00:00:00", "2024-01-01 01:00:00"),
  humidity = c(68.2, 69.0),
  temperature = c(50.4, 48.6),
  pm2.5_alt = c(3.4, 2.0),
  sensor_index = c(1234, 1234),
  stringsAsFactors = FALSE
)
```

```{r, eval=TRUE}
print(history)
```
